---
title: "PCOS_ExploratoryDataAnalysis"
author: "Matt"
date: "2023-01-11"
output: html_document
---


# Introduction 
Incomplete

# Background
Incomplete

# Objective
Utilize the publicly available [Polycystic Ovary Syndrome (PCOS) Dataset found on Kaggle](https://www.kaggle.com/datasets/prasoonkottarathil/polycystic-ovary-syndrome-pcos?resource=download) to propose a model to diagnose PCOS using techniques and concepts covered in MEDI 504B at the University of British Columbia (UBC).  

# Methods
Incomplete

# Results
Incomplete

## Exploratory Data Analysis

### Load Required Libraries
```{r}

library(tidyverse)
library(readxl)
library(DataExplorer)
library(janitor)

```


### Import Dataset
```{r, message=FALSE}
PCOSData <- read_xlsx("Week 1/Lab/PCOS_data_without_infertility.xlsx", 
                                        sheet = "Full_new")
```


### Look at Data
```{r}

#PCOSData
PCOSData # look at the full dataset in raw form
dim(PCOSData) # dataset dimensions
n_distinct(PCOSData$`Patient File No.`) # confirming that there are no repeated patient file numbers

```

We have a tibble, `PCOSData`, containing data collected from 541 unique individuals. This tibble contains data relevant to infertility, each individuals Polycystic Ovary Syndrome Status and much more. A total of 45 variables are present. 

With a quick scan, `PCOSData` looks very complete, other than the column named `...45`. Let's do a more formal check to understand missing data counts...


### Checking Data Completeness

```{r, fig.height=7, fig.width=8}

plot_missing(PCOSData)

```

Great. We can see that, as expected based on initial observations, the `...45` column is missing lots of data (nearly empty!). We can omit this in the next code chunk using select(). There is also very minimally missing data (1 observation/541, 0.18%) in the columns `Fast food (Y/N)` and `Marraige Status (Yrs)` (only 0.18% missing). I'll drop the rows with these missing observations, too. Let's also clean up the variable names and types before going any further.


### Drop Missing Values, Review Variable Types
```{r}
(PCOSDataCLEAN <- PCOSData %>% 
  select(-...45) %>%  #removing the 45th variable/column, as no data was present
  drop_na() %>% #removing remaining rows with N/A's (i.e. the two rows missing fast food and marriage status observations)
   
   #convert categorical variables to factor type
  mutate_at(c("PCOS (Y/N)", "Blood Group", "Pregnant(Y/N)", "Weight gain(Y/N)", "hair growth(Y/N)", 
              "Skin darkening (Y/N)", "Hair loss(Y/N)", "Pimples(Y/N)", 
              "Fast food (Y/N)", "Reg.Exercise(Y/N)"), 
            factor) %>% 
  
  #cleaning up variable names for consistency
  clean_names())

dim(PCOSDataCLEAN)
```

Great. Now we have a tibble that's clean and contains complete observations for 539 unique individuals. All variable types are also relevant. One other thing that would be good to do is re-code the factor levels in the `Blood Type` column, which are currently represented by a number code. I will also correct the number formatting in the `FSH/LH` column by recalculating the ratio.

```{r}
levels(PCOSDataCLEAN$blood_group)
```

Lets recode this... keep in mind that the documentation for the data, found [here](https://www.kaggle.com/datasets/prasoonkottarathil/polycystic-ovary-syndrome-pcos?resource=download&select=PCOS_data_without_infertility.xlsx), specifies the following mappings: 

`A+` = "11", 
`A-` = "12",
`B+` = "13",
`B-` = "14",
`O+` = "15",
`O-` = "16",
`AB+` = "17",
`AB-` = "18"

```{r}
PCOSDataCLEAN <- PCOSDataCLEAN %>% 
  
  #recoding factor levels
  mutate(blood_group = fct_recode(blood_group, 
             `A+` = "11", 
             `A-` = "12",
             `B+` = "13",
             `B-` = "14",
             `O+` = "15",
             `O-` = "16",
             `AB+` = "17",
             `AB-` = "18"))

PCOSDataCLEAN
```

Awesome, this tibble is looking clean. Lets explore some more! Yulia was kind enough to introduce our class to the DataExplorer Package. Let's test it out!

### Using the DataExplorer Package to Explore the Cleaned Data

```{r}
plot_str(PCOSDataCLEAN) #reviewing variable types and dataset dimension

plot_missing(PCOSDataCLEAN) #confirming our data cleaning efforts above were a success

plot_histogram(PCOSDataCLEAN) #visualizing distributions of continuous variables

plot_correlation(PCOSDataCLEAN)

plot_bar(PCOSDataCLEAN) #visualizing categorical variable frequency

# create_report(PCOSDataCLEAN) # a useful collection of all the above features!!!
```

Comments: 


# Supplementary Information
Incomplete

# Data Dictionary
Incomplete


