---
title: "PCOS_ExploratoryDataAnalysis"
author: "Matt"
date: "2023-01-11"
output: html_document
---

### Libraries
```{r}
library(tidyverse)
library(readxl)
library(DataExplorer)
library(janitor)
```

1- Introduction and Background

2- Objectives

3- Methods

4- Results

4a) Exploratory data analysis

5- Supplementary information

5a) Data dictionnary


### Import Datasets

```{r}
PCOSData <- read_xlsx("Week 1/Lab/PCOS_data_without_infertility.xlsx", 
                                        sheet = "Full_new")
```


### Look at Data

```{r}

#PCOSData
PCOSData 
dim(PCOSData)
n_distinct(PCOSData$`Patient File No.`)

```

We have a tibble, `PCOSData`, containing data collected from 541 unique individuals. This tibble contains data relevant to infertility, each individuals Polycystic Ovary Syndrome Status and much more. A total of 45 variables are present. 


With a quick scan, `PCOSData` looks very complete, other than the column named `...45`. Let's do a more formal check to understand missing data counts:

```{r}
plot_missing(PCOSData)
```

Great. We can see that, as expected based on initial observations, the `...45` column is missing lots of data (nearly empty!). We can omit this in the next code chunk using select(). There is also very minimally missing data (1 observation) in the columns `Fast food (Y/N)` and `Marraige Status (Yrs)` (only 0.18% missing). I'll drop the rows with these missing observations. 

Let's also clean up the variable types before going any further.

### Drop Missing Values, Review Variable Types

```{r}
(PCOSDataCLEAN <- PCOSData %>% 
  select(-...45) %>%  #removing the 45th variable/column, as no data was present
  drop_na() %>% #removing remaining rows with N/A's (i.e. the two rows missing fast food and marriage status observations)
   
   #convert categorical variables to factor type
  mutate_at(c("PCOS (Y/N)", "Blood Group", "Pregnant(Y/N)", "Weight gain(Y/N)", "hair growth(Y/N)", 
              "Skin darkening (Y/N)", "Hair loss(Y/N)", "Pimples(Y/N)", 
              "Fast food (Y/N)", "Reg.Exercise(Y/N)"), 
            factor))

dim(PCOSDataCLEAN)
```

Great. Now we have a tibble that's clean and contains complete observations for 539 unique individuals. All variable types are also relevant. One other thing that would be good to do is re-code the factor levels in the `Blood Type` column, which are currently represented by a number code. 

```{r}
levels(PCOSDataCLEAN$`Blood Group`)
```

Lets recode this... keep in mind that the documentation for the data, found here: https://www.kaggle.com/datasets/prasoonkottarathil/polycystic-ovary-syndrome-pcos?resource=download&select=PCOS_data_without_infertility.xlsx, specifies the following mappings: 

`A+` = "11", 
`A-` = "12",
`B+` = "13",
`B-` = "14",
`O+` = "15",
`O-` = "16",
`AB+` = "17",
`AB-` = "18"

```{r}
PCOSDataCLEAN <- PCOSDataCLEAN %>% 
  
  #recoding factor levels
  mutate(`Blood Group` = fct_recode(`Blood Group`, 
             `A+` = "11", 
             `A-` = "12",
             `B+` = "13",
             `B-` = "14",
             `O+` = "15",
             `O-` = "16",
             `AB+` = "17",
             `AB-` = "18"))

PCOSDataCLEAN
```

Awesome, this tibble is looking great. Lets explore some more! Yulia was kind enough to introduce our class to the DataExplorer Package. Let's test it out!



```{r}

```



